### 2020-10-20 submission/manuscript.Rmd
Create Figures 4 and S5

**How does the increased inter-sample variation impact the ability to detect differences between samples?** To test the effect of increased inter-sample variation, I randomly assigned samples to one of two treatment groups. In the first treatment group, communities were randomly sampled from the null distribution as described above. For the second treatment group, I increased the abundance of 10% of the ASVs in the pooled study distribution by 5%. I randomly generated 100 simulated sets of treatment groups and samples. I then tested the ability to detect a difference between the two treatment groups using alpha and beta diversity metrics. The fraction of significant tests was a measurement of the statistical power to detect the difference between the treatment groups. Across all of the metrics I tested, the marine dataset yielded no simulated sets that were statistically significant, which was likely due to the small number of samples in the study (N=7); the marine dataset is not included in the following analysis. Among the remaining datasets, the power to detect a difference in the observed number of ASVs ranged between `r oxford_comma(b_base_pc_sobs, 2)` and between `r oxford_comma(b_base_pc_shannon, 2)` to detect a difference in diversity when using a Wilcox test (**Figure X**). When singleton ASVs were removed, the power to detect a difference in the number of ASVs dropped by between `r oxford_comma(b_pc_sobs, 1)`% and by between `r oxford_comma(b_pc_shannon, 1)`% to detect a difference in diversity. The effect of removing rare ASVs on the number of OTUs and their diversity was similar. A popular alternative approach calculates the dissimilarity between communities based on the abundance of each ASV or OTU. We used the Bray-Curtis dissimilarity index to compare the simulated communities within each dataset and calculated the power to detect differences between the two simulated treatment groups using the analysis of molecular variance (also called PERMANOVA). Without removing rare sequences, the power to detect a difference between the two simulated treatment groups varied between `r oxford_comma(b_base_beta, 2)`. Aside from `r n_unfazed_singletons` datasets, the power to detect differences dropped by between `r oxford_comma(fazed_singletons, 1)`% when singletons were removed. However, when ASVs that occurred 10 or fewer times were removed from each sample, the power to detect differences dropped by `r oxford_comma(fazed_tentons, 1)`%; similar results were observed when ASVs were clustered into OTUs. Removing rare ASVs reduced the ability to detect simulated treatment effects using metrics commonly used to compare microbial communities.

>>>>

**How does the increased inter-sample variation impact the ability to detect differences between samples?** To test the effect of increased inter-sample variation, I randomly assigned samples to one of two treatment groups. In the first treatment group, communities were randomly sampled from the null distribution as described above. For the second treatment group, I increased the abundance of 10% of the ASVs in the pooled study distribution by 5%. I randomly generated 100 simulated sets of treatment groups and samples. I then tested the ability to detect a difference between the two treatment groups using alpha and beta diversity metrics. The fraction of significant tests was a measurement of the statistical power to detect the difference between the treatment groups.

This doesn't seem to agree with the figure...
Across all of the metrics I tested, the marine dataset yielded no simulated sets that were statistically significant, which was likely due to the small number of samples in the study (N=7); the marine dataset is not included in the following analysis.

Among the remaining datasets, the power to detect a difference in the observed number of ASVs ranged between `r oxford_comma(b_base_pc_sobs, 2)` and between `r oxford_comma(b_base_pc_shannon, 2)` to detect a difference in diversity when using a Wilcox test (Figure 4A). When singleton ASVs were removed, the power to detect a difference in the diversity of ASVs dropped by between `r oxford_comma(b_pc_sobs, 1)`% and by between `r oxford_comma(b_pc_shannon, 1)`% (Figure 4B). The effect of removing rare ASVs on the number of OTUs and their diversity was similar (Figure S5AB). A popular alternative approach calculates the dissimilarity between communities based on the abundance of each ASV or OTU. We used the Bray-Curtis dissimilarity index to compare the simulated communities within each dataset and calculated the power to detect differences between the two simulated treatment groups using the analysis of molecular variance (also called PERMANOVA) (Figure 4C and S5C). Without removing rare sequences, the power to detect a difference between the two simulated treatment groups varied between `r oxford_comma(b_base_beta, 2)`. Aside from `r n_unfazed_singletons` datasets, the power to detect differences dropped by between `r oxford_comma(fazed_singletons, 1)`% when singletons were removed. However, when ASVs that occurred 10 or fewer times were removed from each sample, the power to detect differences dropped by `r oxford_comma(fazed_tentons, 1)`%; similar results were observed when ASVs were clustered into OTUs. Removing rare ASVs reduced the ability to detect simulated treatment effects using metrics commonly used to compare microbial communities.
